LCS Tool: A Computational Platform for Lagrangian Coherent Structures

### 2014-11-06 main.tex
Rename strainline to shrinkline

Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrinklines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_1(r),
\label{eq:strainline}
\end{equation}
and attracting LCSs (\emph{stretchlines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_2(r).
\label{eq:stretchline}
\end{equation}

@@

This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:strainline,eq:stretchline}. \Cref{t:LCS Tool overview} gives an overview of the steps, functions and variable names used in LCS Tool to analyse a flow.

@@

As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:strainline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs). By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$. We identify the strongest repelling shrinklines as those crossing a local maximum of the $\lambda_2(x_0)$ field. Similarly, we identify the strongest attracting stretchlines as those crossing a local minimum of the $\lambda_1(x_0)$ field. These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.

@@

\begin{table}
\begin{enumerate}
\item Define a local maximization distance.
\item Find all points of the main grid that are local maxima of $\lambda_2$ within a circle whose radius is the local  maximization distance.
\item Define a maximum shrinkline length
\item Integrate a shrinkline forward and backward according to \cref{eq:strainline} and using the largest $\lambda_2$ local maximum as the initial position. Integrate until the shrinkline has attained the maximum shrinkline length, or until it has reached the domain boundary.
\item Flag any remaining local maxima of $\lambda_2$ within the maximization distance of the shrinkline as ineligible initial positions for subsequent shrinklines
\label{i:flag maxima}
\item Continue integrating shrinklines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
\item Remove all shrinkline segments within elliptic LCSs.
\end{enumerate}
\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
\label{t:Hyperbolic LCS algorithm}
\end{table}

@@

\Cref{f:double gyre lambda hyperbolic LCS} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution. \Cref{l:double gyre lambda strain stretch LCS} lists the corresponding LCS Tool commands. The maximal length of shrinklines and stretchlines, \lstinline!strainlineMaxLength! and \lstinline!stretchlineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.

@@

% Local Variables:
% ispell-dictionary: "en_GB"
% End:
% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected strainlines minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry zonal

>>>>

Solutions to this variational problem turn out to be orbits of the $\xi_1$ or $\xi_2$ eigenvector field. Repelling LCSs (\emph{shrinklines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_1(r),
\label{eq:shrinkline}
\end{equation}
and attracting LCSs (\emph{stretchlines}) are obtained as trajectories of the differential equation
\begin{equation}
r' = \xi_2(r).
\label{eq:stretchline}
\end{equation}

@@

This section describes stepwise the numerical implementation of geodesic detection of elliptic and hyperbolic LCSs based on \cref{eq:etafields,eq:shrinkline,eq:stretchline}. \Cref{t:LCS Tool overview} gives an overview of the steps, functions and variable names used in LCS Tool to analyse a flow.

@@

As discussed in \cref{sec:Hyperbolic LCSs}, positions of hyperbolic LCSs at time $t_0$ are found as the strongest repelling orbits of the vector field \cref{eq:shrinkline} (repelling LCSs), and strongest attracting orbits of the vector field \cref{eq:stretchline} (attracting LCSs). By repulsion and attraction we mean a property of the LCS (as an evolving material line) under the flow map $F_{t_0}^t$. We identify the strongest repelling shrinklines as those crossing a local maximum of the $\lambda_2(x_0)$ field. Similarly, we identify the strongest attracting stretchlines as those crossing a local minimum of the $\lambda_1(x_0)$ field. These local maxima and minima of the appropriate $\lambda_i(x_0)$ eigenvalue field can be thought of as the extensions of the concept of saddle points to the present finite-time, temporally aperiodic flow setting.

@@

\begin{table}
\begin{enumerate}
\item Define a local maximization distance.
\item Find all points of the main grid that are local maxima of $\lambda_2$ within a circle whose radius is the local  maximization distance.
\item Define a maximum shrinkline length
\item Integrate a shrinkline forward and backward according to \cref{eq:shrinkline} and using the largest $\lambda_2$ local maximum as the initial position. Integrate until the shrinkline has attained the maximum shrinkline length, or until it has reached the domain boundary.
\item Flag any remaining local maxima of $\lambda_2$ within the maximization distance of the shrinkline as ineligible initial positions for subsequent shrinklines
\label{i:flag maxima}
\item Continue integrating shrinklines using local maxima of $\lambda_2$ as initial positions until no eligible local maxima of $\lambda_2$ remain.
\item Remove all shrinkline segments within elliptic LCSs.
\end{enumerate}
\caption{Algorithm to calculate initial positions of repelling LCSs at time $t_0$. The algorithm for attracting LCSs is similar.}
\label{t:Hyperbolic LCS algorithm}
\end{table}

@@

\Cref{f:double gyre lambda hyperbolic LCS} shows the complete picture of elliptic and hyperbolic LCSs of the double gyre at this resolution. \Cref{l:double gyre lambda strain stretch LCS} lists the corresponding LCS Tool commands. The maximal length of shrinklines and stretchlines, \lstinline!shrinklineMaxLength! and \lstinline!stretchlineMaxLength!, is set to $20$, a multiple of the domain size, since the hyperbolic LCSs may wind around vortices several times.

@@

% Local Variables:
% ispell-dictionary: "en_GB"
% End:
% LocalWords:  stretchline stretchlines incompressible gyre eigenvector eigenvectors incompressibility dataset datasets LCS LCSs Lyapunov Bickley Rossby timespan variational adaptively vortices parametrise advected minima extremal FTLE MATLAB maxima geostrophic shrinkline shrinklines Poincare altimetry zonal
